Article Title

### 2014-01-16 plos.tex
removed illustration in Introduction

\DeclareMathOperator{\Tr}{tr}
\newcommand{\sq}[1]{\lq#1\rq}
\newcommand{\mcond}{\,\middle\vert\,}
\newcommand{\cond}{\,\vert\,}
\newcommand{\figref}[2]{Fig.\;\ref{fig:#1}\,#2}
\newcommand{\loss}[1]{\mathcal L\left(#1\right)} 
\newcommand{\eloss}[1]{\mathcal L_0\left(#1\right)}
\newcommand{\T}{{\sf T}}
\newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value
\newcommand{\TODO}[1]{\emph{\small\color{blue}$\langle\langle$#1$\rangle\rangle$}}
\newcommand*\dif{\mathop{}\,d}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\rank}{rank}

@@

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization, \emph{i.e.}\;by imposing a structure on the estimate.  The amount of improvement depends on how closely the assumed structure represents dependencies in the data. Therefore, the selection of the most efficient correlation matrix estimator for a given neural circuit must be determined empirically.  Furthermore, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.

We sought the most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix should be most efficiently represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.  Indeed, in cross-validation tests, the covariance matrix estimator with such structure consistently outperformed other regularized estimators. The connectivity patterns inferred from the sparse component of the estimate manifested a strong relationship to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distance and the difference in orientation preference whereas the negative \sq{inhibitory} connections were less selective.  Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than sample correlations.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations in the activity of neural populations have proven useful as descriptions of the functional connectivity in the embedding circuit.  Correlations may arise from various underlying processes that reflect differently on the structure of the correlation matrix:  Common network fluctuations or external inputs are well approximated by low-rank representations whereas linear effects between specific pairs of neurons are well approximated by fitting their pairwise partial correlations.  What kinds of correlation effects dominate the population activity in cortical microcircuits? To answer this question, we imposed various correlation structures on empirical estimates of neural correlation matrices and evaluated which of them produced greatest estimation improvement.  In fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the greatest improvement was produced by decomposing the correlation matrix as a sparse network of partial correlations and a low-rank component.  Since it leads to the best estimates, we propose that this structure provides a better picture of the functional connectivity than that obtained from sample correlations. As an application of this approach, we analyzed how the inferred connectivity related to distances between cells and to differences in their preferred orientations and found basic resemblance to results from studies of synaptic connectivity.
\section*{Introduction}
\paragraph{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the functional connectivity in neural circuits \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010,Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.

However, neural correlations do not submit to ready or unambiguous mechanistic interpretation.  Theoretical studies and simulations have shown that neural correlations on various temporal scales may arise from combinations of multiple mechanisms including direct synaptic interactions, common inputs or correlated inputs, chains of multiple synaptic connections, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}.

\paragraph{Structure  of correlation matrices}
Modern multineuronal recordings allow measuring entire correlation matrices from increasingly large and dense populations of highly interconnected neurons. Multivariate analysis techniques based on large correlation matrices provide information that could not be extracted from an equivalent number of isolated measurements of pairwise correlations.  For example, when the correlation matrix is approximated in a form with a reduced number of variables, this \emph{reduced representation} can isolate an important aspect of the functional connectivity in neural circuits and possibly point to underlying physiological mechanisms and computational principles.

One reduced representation of the correlation matrix is its low-rank approximation.  Low-rank approximations are particularly suitable for capturing common fluctuations and common external inputs and can be extracted using principal component analysis, independent component analysis, factor analysis, or  \cite{Chapin:1999, Peyrache:2010, Lopes:2011, Lopes:2013}.

Another reduced representation of the correlation matrix can be obtained by \emph{covariance selection}, \emph{i.e.}\;by selecting a subset of the pairwise \emph{partial correlations} that are sufficient to reproduce the entire correlation matrix \cite{Dempster:1972, Friedman:2008}. The partial correlation between a pair of neurons is a measure of their linear association after the effects of the rest of the network have been taken into account \cite{Anderson:2003}. Partial correlations are particularly meaningful when all effects between the variables are nearly linear and a large fraction of the variables are observed; in this case, partial correlations approximate conditional dependency.  A lack of conditional dependency between two neurons suggests a lack of direct interaction between them. Networks of partial correlations (sometimes called \emph{association networks}) have been used to infer gene interaction networks \cite{Schafer:2005, Peng:2009} and interactions between brain regions in brain imaging studies \cite{Varoquaux:2012, Ryali:2012}.

\paragraph{Improving estimation through regularization}
Imposing a structure on an estimate of the covariance matrix can improve its estimation performance.

Estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.  This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}.

In neuroscience studies, the true correlation matrix is usually estimated by the \emph{sample correlation matrix}, which is rarely distinguished from the true correlation matrix.  The sample correlation matrix is unbiased but, because it has many free parameters, it is highly sensitive to sampling noise. As a result, on average, the sample correlation matrix is far from the true correlation matrix.  In contrast, reduced estimates are typically less susceptible to sampling noise although can introduce their respective biases.

In statistics, the technique of deliberately imposing a structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.    To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.  The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.  Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely. Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).  Yet others can do both target selection and shrinkage toward the target.

Although regularized covariance matrix estimates have become commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.  
\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972, Friedman:2008}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix (inverse of the covariance matrix) while setting the rest to zero.  Zeros in the off-diagonal elements of the precision matrix indicate zero partial correlations between the corresponding pair of neurons.   Therefore, covariance selection amounts to finding an approximation of the covariance matrix that maximizes the number of neuron pairs that are linearly independent of each other when conditioned on the activity of all the other observed cells.

We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients.

Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Conceptual questions and basic approach}
What structures of correlation matrices best describe the multineuronal activity in specific circuits and in specific brain states?  More specifically, are correlations in the visual cortex during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?

To address these questions, we evaluated five covariance matrix estimators based on different assumptions about the underlying correlation structure:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix. 
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis.
\item[$C_{\sf sparse}$] --  sparse partial correlations, \emph{i.e.}\;covariance selection.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations with several inferred latent units.
\end{description}

First, we demonstrated that, in simulations with known true correlation structures, estimators that imposed a matching structure were generally more efficient than the other estimators.  The estimators’ relative performance was evaluated by cross-validation.   The logic of this study relies on the converse of this relationship: the covariance matrix estimator that is found to be most efficient on empirical data is proposed to provide a better picture of the correlation structure than other estimators.  Caveats for this type of inference are reviewed in the Discussion.

Therefore, we performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with fast 3D random-access two-photon imaging of calcium signals. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units.  We examined the structure of the sparse partial correlations inferred by the estimate with respect to the neurons’ geometrical positions and preferred orientations and found that stronger dependencies than those of sample correlations. In Discussion, we outline possible approaches to further corroborate the relationship between the structure of the functional connectivity revealed by such methods to the anatomical organization of the circuit.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The icovariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates. In the case of noise correlations, $\mu$ depends on the stimulus condition.  Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.  In the case of noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.

Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}\label{eq:precision}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
Similarly, the matrix of partial correlations $P$ is computed by normalizing the the negative \emph{precision matrix} $C^{-1}$ (inverse of the covariance matrix):
\begin{equation}\label{eq:partial}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. Clearly, off-diagonal zeros in the precision matrix correspond to zero partial correlations.

We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, their targets estimates can be described by distinct families of graphical models (\figref{02}{row 1}).

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Estimators whose low-dimensional regularization targets can represent the structure of the true covariance matrix outperform other estimators.}
        {\bf Row 1.} Graphical representations of four types of low-dimensional structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
        In the \sq{independent model} ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
        In \sq{latent factors} ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
        In \sq{sparse interactions} ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
        In \sq{sparse+latent} ({\bf D}), correlations arise through direct linear interactions in a sparse subset of observed neurons and through interactions with common latent units. 
        {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
        The factor model ({\bf B}) has three latent units. 
        The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
        The \sq{sparse+latent} matrix has one latent unit and its direct interactions are 78\% sparse.
        {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
        {\bf Row 4.} Mean \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.
        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator. The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:02}
\end{FPfigure}

For estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal contains  variance estimates. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure of target $D$ favors the absence of linear associations between the activity of observed neurons (\figref{02}{A}).  If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ should outperform the other estimators.

@@

Estimator $C_{\sf sparse}$ is based on the assumption that correlations result from direct linear associations between pairs of observed neurons and that only a fraction of neuron pairs interact (\figref{02}{C}).  This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero. The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.  The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.

>>>>

\DeclareMathOperator{\Tr}{tr}
\newcommand{\sq}[1]{\lq#1\rq}
\newcommand{\mcond}{\,\middle\vert\,}
\newcommand{\cond}{\,\vert\,}
\newcommand{\figref}[2]{Fig.\;\ref{fig:#1}\,#2}
\newcommand{\loss}[1]{\mathcal L\left(#1\right)} 
\newcommand{\eloss}[1]{\mathcal L_0\left(#1\right)}
\newcommand{\T}{{\sf T}}
\newcommand{\E}[2][]{\mathbb E_{#1}\left[ #2\right]}    % expected value

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\rank}{rank}

@@

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the organization of neural circuits.  However, estimating and interpreting large correlation matrices is challenging.  Estimation can be improved by regularization, \emph{i.e.}\;by imposing a structure on the estimate.  The amount of improvement depends on how closely the assumed structure represents dependencies in the data. Therefore, the selection of the most efficient correlation matrix estimator for a given neural circuit must be determined empirically.  Importantly, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.

We sought statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix should be best represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.  Indeed, in cross-validation tests, the covariance matrix estimator with such a structure consistently outperformed other regularized estimators. The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.  Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.

It is now possible to record the spiking activity of hundreds of neurons at the same time.  A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.  We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.  Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.  In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (\sq{interactions}) combined with a low-rank component. The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.  We propose that the composition of the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
\section*{Introduction}
\emph{Functional connectivity} is a statistical description of \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.  Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons or in the form of joint coactivation patterns.  Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.  To remove the confounding effects of the stimulus, the average stimulus responses are subtracted or otherwise accounted for. Although functional connectivity is only a phenomenological description, it can serve as a tool to generate hypotheses about how population activity relates to the anatomical architecture of neural circuits and informs about how information is encoded and processed at the population level.

Pearson correlations between the spiking activity of pairs of neurons, are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  In particular, \emph{noise correlations}, \emph{i.e.}\;the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the anatomical connectivity \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010} among others.

Neural correlations do not come with ready or unambiguous mechanistic interpretations. They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}. But multineuronal recordings do provide more information than available from an equivalent number of pairs of cells considered separately. For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients. In contrast, the off-diagonal elements of the precision matrix (inverse of the covariance) constitute scaled partial correlations between neuron pairs.  The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.  For this reason, partial correlations have been proposed to more closely reflect the direct effects between components of complex systems and have been used to describe gene interactions \cite{Schafer:2005, Peng:2009} and the interactions between brain regions \cite{Varoquaux:2012, Ryali:2012}.

However, estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.  This mismatch leads to an increase in spurious correlations, overestimation of common activity (\emph{i.e.}\;overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}. 
The \emph{sample correlation matrix}, which is an unbiased estimate, is often used to estimate the true correlation matrix (ref). However, it has many free parameters, and is thus highly sensitive to sampling noise. As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.

In statistics, the technique of deliberately imposing a structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.    To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.  The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.  Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely. Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).  Yet others do both target selection and shrinkage.

Although regularized covariance matrix estimation is commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.  All prior empirical findings from neural correlations have relied on sample correlations. Improved estimation is not the only or even the main benefit of regularization.  When we select the best performing estimator among many regularized estimators, we learn something about the system itself:  The structure imposed by the best estimator represents the underlying dependencies in the system more closely than the structures imposed by the other estimators.  We can then examine this structure as a more effective description of the functional connectivity in the circuit than the sample correlation matrix.

%\paragraph{Conceptual questions and basic approach}
What structures of correlation matrices best describe the multineuronal activity in specific circuits and in specific brain states?  More specifically, are correlations in the visual cortex during visual stimulation best explained by common fluctuations or by local interactions within the recorded microcircuit?

To address these questions, in addition to the sample covariance matrix, we evaluated four regularized covariance matrix estimators that imposed different kinds of structure on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix. 
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis.
\item[$C_{\sf sparse}$] --  sparse partial correlations, \emph{i.e.}\;covariance selection.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations with several inferred latent units.
\end{description}

First, we demonstrated that, in simulations with known correlation structures, estimators that imposed a matching structure were generally more efficient than other estimators. We therefore reasoned as follows: The most efficient covariance matrix estimator is likely to provide a better picture of the correlation structure than less efficient estimators (See Discussion).

We therefore performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units with partial correlations with the observed cells. We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive the inferred partial correlations had a large fraction of negative values possibly reflecting inhibitory circuitry.  Moreover, we found that the inferred positive partial correlations exhibited stronger dependencies with respect to physical distance and differences in preferred orientations than noise correlations. In contrast, the inferred negative partial correlations were less selective.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The covariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates.  Usually, neural covariance matrices have been estimated as the \emph{sample covariance matrix} $C_{\sf 0}$ calculated from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.  In the case of noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.

Given a covariance matrix estimate $C$ (not necessarily the sample covariance $C_{\sf 0})$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}\label{eq:precision}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2},
\end{equation}
where $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$ (inverse of the covariance matrix):
\begin{equation}\label{eq:partial}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}

We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, their targets estimates can be described by distinct families of graphical models (\figref{02}{\,Row}{row 1}).

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Estimators whose regularization targets represent the structure of the true covariance matrix outperform other estimators.}
        {\bf Row 1.} Graphical representations of the regularization targets in the four regularized covariance matrix estimators used in this study,   structures of interactions between observed neurons (green spheres) and latent units (light-shaded spheres).
        In $C_{\sf diag}$, the regularization target is  \sq{diagonal} ({\bf A}), observed neurons exert no linear effects on one another neither directly nor through interactions with common latent units. 
        In \sq{latent factors} ({\bf B}), the correlated activity of observed cells is driven by several latent units. 
        In \sq{sparse interactions} ({\bf C}), the correlation matrix is defined by a set of linear interactions between observed neurons. 
        In \sq{sparse+latent} ({\bf D}), correlations arise through direct linear interactions in a sparse subset of observed neurons and through interactions with common latent units. 
        {\bf Row 2.} Examples of $50\times 50$ correlation matrices corresponding to each type of low-dimensional structure. 
        The factor model ({\bf B}) has three latent units. 
        The partial correlation matrix of the sparse model ({\bf C}) is 73\% sparse.
        The \sq{sparse+latent} matrix has one latent unit and its direct interactions are 78\% sparse.
        {\bf Row 3.} Examples of sample correlation matrices calculated from samples of 1000 observations taken from simulated random processes with corresponding correlation matrices from row 2.
        {\bf Row 4.} Mean \emph{excess loss} (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.
        {\bf Row 5.} Mean \emph{cross-validation loss} (Eq.~\ref{eq:rel-cv-loss}) of covariance estimators with respect to the matching estimator. The values are relative to the validation loss of the estimator that matches the low-dimensional structure of the true covariance matrix. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:02}
\end{FPfigure}

For estimator $C_{\sf diag}$, the target estimate is the diagonal matrix $D$ whose diagonal elements contain the variance estimates. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure of target $D$ favors the absence of linear associations between the activity of observed neurons (\figref{02}{A Row 1}).  If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ should outperform the other estimators.

@@

Estimator $C_{\sf sparse}$ favors correlations that arise from direct linear effects between pairs of observed neurons when only a fraction of neuron pairs interact (\figref{02}{C}).  This assumption is enforced by fitting only a subset of the off-diagonal elements of the precision matrix while setting the rest to zero. The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.  The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.
